---
title: "Simple examples of principal graphs"
author: "Luca Albergante (luca.albergante@gmail.com)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette present a few simple examples describing the usage of the package RPrinGraph. For a more deailed description of the package and for various instruction on installation refer to a different vignette.

## Pircipal elastic circles

This may be necessary for Mac systems:

```{r}
dyn.load('/Library/Java/JavaVirtualMachines/jdk1.7.0_79.jdk/Contents/Home/jre/lib/server/libjvm.dylib')
```

```{r}
library(rpcurve)

Data <- simple_circle

Results <- computeElasticPrincipalGraph(Data = Data, NumNodes = 40, Method = 'CircleConfiguration')

```


```{r, fig.show='hold'}
plotMSDEnergyPlot(Results, Main = "Pincipal Circle", Cex.Main = 1)
```

```{r, fig.show='hold'}
accuracyComplexityPlot(Results, Main = "Pincipal Circle", Cex.Main = 1, Mode = 5)
```

```{r, fig.show='hold'}
accuracyComplexityPlot(Results, Main = "Pincipal Circle", Cex.Main = 1, Xlims = c(.9, 1))
```

```{r, fig.show='hold'}
accuracyComplexityPlot(Results, Main = "Pincipal Circle", Cex.Main = 1, Xlims = c(.97, .98))
```


```{r, fig.show='hold'}
plotData2D(Data = simple_circle, PrintGraph = Results,
           GroupsLab = rep(1, nrow(simple_circle)), Xlab = "Dimension 1", Ylab = "Dimension 2")
```


```{r, fig.show='hold'}
plotData2D(Data = simple_circle, PrintGraph = Results, Plot.ly = TRUE,
           GroupsLab = rep(1, nrow(simple_circle)), Xlab = "Dimension 1", Ylab = "Dimension 2")
```


```{r, fig.show='hold'}
plotData3D(Data = simple_circle, PrintGraph = Results, Plot.ly = TRUE,
           GroupsLab = rep(1, nrow(simple_circle)))
```

```{r, fig.show='hold'}
library("rgl")
open3d()
plotData3D(Data = simple_circle, PrintGraph = Results, Plot.ly = FALSE,
           GroupsLab = rep(1, nrow(simple_circle)))
```




```{r, fig.show='hold'}
Net <- ConstructGraph(Results = Results, DirectionMat = NULL, Thr = 0.05)
TaxonList <- getTaxonMap(Graph = makeGraph(Results), Data = Data)

InfoData <- plotPieNet(Results = Results, Data = simple_circle, Categories = rep(1, nrow(simple_circle)),
           Graph = Net, TaxonList = TaxonList, LayOut = 'circle', Main = "Pincipal Circle")
```


## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
