---
title: "Using the package"
author: "Luca Albergante (luca.albergante@gmail.com)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette present a few simple examples describing the usage of the package RPrinGraph. For a more deailed description of the package and for various instruction on installation refer to a different vignette.


## An essential overview of Pircipal elastic circles


## rpgraph prerequisite and instakllation

dyn.load('/Library/Java/JavaVirtualMachines/jdk1.7.0_79.jdk/Contents/Home/jre/lib/server/libjvm.dylib')





This may be necessary for Mac systems:

## Workaround for common problems




## Example 1 - Circle


```{r}
library(rpcurve)

Data <- simple_circle

Results <- computeElasticPrincipalGraph(Data = Data, NumNodes = 40, Method = 'CircleConfiguration')

```


```{r, fig.show='hold'}
plotMSDEnergyPlot(Results, Main = "Pincipal Circle", Cex.Main = 1)
```

```{r, fig.show='hold'}
accuracyComplexityPlot(Results, Main = "Pincipal Circle", Cex.Main = 1, Mode = 5)
```

```{r, fig.show='hold'}
accuracyComplexityPlot(Results, Main = "Pincipal Circle", Cex.Main = 1, Xlims = c(.9, 1))
```

```{r, fig.show='hold'}
accuracyComplexityPlot(Results, Main = "Pincipal Circle", Cex.Main = 1, Xlims = c(.97, .98))
```


```{r, fig.show='hold'}
plotData2D(Data = simple_circle, PrintGraph = Results,
           GroupsLab = rep(1, nrow(simple_circle)), Xlab = "Dimension 1", Ylab = "Dimension 2")
```


```{r, fig.show='hold'}
plotData2D(Data = simple_circle, PrintGraph = Results, Plot.ly = TRUE,
           GroupsLab = rep(1, nrow(simple_circle)), Xlab = "Dimension 1", Ylab = "Dimension 2")
```


```{r, fig.show='hold'}
plotData3D(Data = simple_circle, PrintGraph = Results, Plot.ly = TRUE,
           GroupsLab = rep(1, nrow(simple_circle)))
```

```{r, fig.show='hold'}
library("rgl")
open3d()
plotData3D(Data = simple_circle, PrintGraph = Results, Plot.ly = FALSE,
           GroupsLab = rep(1, nrow(simple_circle)), NodeSizeMult = 0.05)
```




```{r, fig.show='hold'}
Net <- ConstructGraph(Results = Results, DirectionMat = NULL, Thr = 0.05)
TaxonList <- getTaxonMap(Results = Results, Data = Data)

InfoData <- plotPieNet(Results = Results, Data = simple_circle, Categories = rep(1, nrow(simple_circle)),
           Graph = Net, TaxonList = TaxonList, LayOut = 'circle', Main = "Pincipal Circle")
```


## Example 2 - Curve



## Example 3 - tree



## References
