---
title: "Modifying contruction parameters for the principal graphs"
author: "Luca Albergante (luca.albergante@gmail.com)"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

This document describes how to vary the internal parameters used by the java library during the construction of the principal graph.

## Building different principal trees

The construction of principal graphs is controlled by the two parameters **EP**, which indicates the coefficient of elastic stretching of the graph, and **RP**, which indicates the penalty coefficient for the deviation from harmonicity. By default EP is equal to 0.01 and RP to 0.1. However it is possible to change these parameters direclty when building the graph. Changing these values will affect how the princial graph is fitted to the data.

```{r}
library(rpgraph)
Results <- computeElasticPrincipalGraph(Data = simple_tree, NumNodes = 20, Method = 'DefaultPrincipalTreeConfiguration')
```

```{r, fig.height=7, fig.width=7}
plotData2D(Data = simple_tree, PrintGraph = Results[[1]] ,
           GroupsLab = rep(1, nrow(simple_tree)),
           Main = "Pincipal tree (EP = .01 and RP = .1)",
           Xlab = "Dimension 1", Ylab = "Dimension 2")
```




```{r}
Results <- computeElasticPrincipalGraph(Data = simple_tree, NumNodes = 20, Method = 'DefaultPrincipalTreeConfiguration', EP = .1)
```


```{r, fig.height=7, fig.width=7}
plotData2D(Data = simple_tree, PrintGraph = Results[[1]] ,
           GroupsLab = rep(1, nrow(simple_tree)),
           Main = "Pincipal tree (EP = .1 and RP = .1)",
           Xlab = "Dimension 1", Ylab = "Dimension 2")
```



```{r}
Results <- computeElasticPrincipalGraph(Data = simple_tree, NumNodes = 20, Method = 'DefaultPrincipalTreeConfiguration', RP = .01)
```


```{r, fig.height=7, fig.width=7}
plotData2D(Data = simple_tree, PrintGraph = Results[[1]] ,
           GroupsLab = rep(1, nrow(simple_tree)),
           Main = "Pincipal tree (EP = .01 and RP = .01)",
           Xlab = "Dimension 1", Ylab = "Dimension 2")
```






```{r}
Results <- computeElasticPrincipalGraph(Data = simple_tree, NumNodes = 20, Method = 'DefaultPrincipalTreeConfiguration', EP = 1, RP = .001)
```


```{r, fig.height=7, fig.width=7}
plotData2D(Data = simple_tree, PrintGraph = Results[[1]] ,
           GroupsLab = rep(1, nrow(simple_tree)),
           Main = "Pincipal tree (EP = 1 and RP = .001)",
           Xlab = "Dimension 1", Ylab = "Dimension 2")
```





