---
title: "Using rpgraph to combine differnt topologies"
author: "Luca Albergante (luca.albergante@gmail.com)"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

This document describes how to combine different graph grammars to produce advanced topologies

## An *hairy circle*

The `simple_circle` dataset included in the package describes points placed on a three dimensional circle. To produce an circle with additional brances we stard by producing a circle

```{r}
library(rpgraph)
PCircle <- computeElasticPrincipalGraph(Data = simple_circle, NumNodes = 20, Method = 'CircleConfiguration')
```

```{r, fig.height=7, fig.width=7}

plotData2D(Data = simple_circle, PrintGraph = PCircle[[1]] ,
           GroupsLab = rep(1, nrow(simple_tree)),
           Main = "Pincipal circle",
           Xlab = "Dimension 1", Ylab = "Dimension 2")

```

It is then possible to add branches by contructing a tree starting from the circle obtained

```{r, fig.height=10, fig.width=10}
HairyCircle <- computeElasticPrincipalGraph(Data = simple_circle, NumNodes = 40,
                                            Method = 'DefaultPrincipalTreeConfiguration',
                                            NodesPositions = PCircle[[1]]$Nodes, Edges = PCircle[[1]]$Edges)
```


```{r, fig.height=7, fig.width=7}

plotData2D(Data = simple_circle, PrintGraph = HairyCircle[[1]] ,
           GroupsLab = rep(1, nrow(simple_tree)),
           Main = "An hairy circle",
           Xlab = "Dimension 1", Ylab = "Dimension 2")

```










## An *lollipop*

Is it also possible to construct a circle with only one branching point. However, this requires a more advanced approach. We will start by adding a strong branch to the `simple_circle` dataset and by getting the initial principal circle.

```{r}
library(rpgraph)

Data <- rbind(simple_circle,
      cbind(seq(1,2,.1) + rnorm(11, sd=0.1),
      seq(1,2,.1) + rnorm(11, sd=0.1),
      rnorm(11, sd=0.1))
)


PCircle <- computeElasticPrincipalGraph(Data = Data, NumNodes = 20, Method = 'CircleConfiguration')
```

we will now apply the tree grammar une step at the time untill we get a single branching points

```{r}
NBranches <- 0

tGraph <- PCircle

while(NBranches != 1){
  tGraph <- computeElasticPrincipalGraph(Data = Data, NumNodes = nrow(tGraph[[1]]$Nodes) + 1,
                                         Method = 'DefaultPrincipalTreeConfiguration',
                                         NodesPositions = tGraph[[1]]$Nodes, Edges = tGraph[[1]]$Edges)
  
  Br <- strsplit(as.character(tGraph[[1]]$Report$BARCODE), "||", fixed = TRUE)[[1]][1]
  NBranches <- sum(as.integer(strsplit(Br, "|", fixed = TRUE)[[1]]))
}

```

We now has a circle with a single branching point. 

```{r, fig.height=7, fig.width=7}
plotData2D(Data = Data, PrintGraph = tGraph[[1]] ,
           GroupsLab = rep(1, nrow(simple_circle)),
           Main = "A basic lollipop",
           Xlab = "Dimension 1", Ylab = "Dimension 2")
```

To avoid adding new branching points, we need to use the principal curve grammar from now on

```{r}
Lollipop <- computeElasticPrincipalGraph(Data = Data, NumNodes = 40,
                                         Method = 'CurveConfiguration',
                                         NodesPositions = tGraph[[1]]$Nodes, Edges = tGraph[[1]]$Edges)
```


```{r, fig.height=7, fig.width=7}
plotData2D(Data = Data, PrintGraph = Lollipop[[1]] ,
           GroupsLab = rep(1, nrow(simple_circle)),
           Main = "A lollipop",
           Xlab = "Dimension 1", Ylab = "Dimension 2")
```



## An tree with a maximum number of branches

Using an approch similar to whas has been done before, it is possible to create a principal tre with a fixed number of branches. Assumme for example that we wat to fit a principal tree to the `simple_tree` dataset with only 3 branches. To do that we start by constructing a tree with only three brances

```{r}
tGraph <- computeElasticPrincipalGraph(Data = simple_tree, NumNodes = 3,
                                       Method = 'DefaultPrincipalTreeConfiguration')

NBranches <- unlist(strsplit(as.character(tGraph[[1]]$Report$BARCODE), "||", fixed = TRUE))[1]

while(NBranches < 3){
  tGraph <- computeElasticPrincipalGraph(Data = simple_tree, NumNodes = nrow(tGraph[[1]]$Nodes) + 1,
                                         Method = 'DefaultPrincipalTreeConfiguration',
                                         NodesPositions = tGraph[[1]]$Nodes, Edges = tGraph[[1]]$Edges)
  
  Br <- strsplit(as.character(tGraph[[1]]$Report$BARCODE), "||", fixed = TRUE)[[1]][1]
  NBranches <- sum(as.integer(strsplit(Br, "|", fixed = TRUE)[[1]]))
}

```


```{r, fig.height=7, fig.width=7}
plotData2D(Data = simple_tree, PrintGraph = tGraph[[1]] ,
           GroupsLab = rep(1, nrow(simple_tree)),
           Main = "A minimal tree with 3 branches",
           Xlab = "Dimension 1", Ylab = "Dimension 2")
```

Now the tree can be extended by using the curve topology

```{r}
tGraph <- computeElasticPrincipalGraph(Data = simple_tree, NumNodes = 50,
                                         Method = 'CurveConfiguration',
                                         NodesPositions = tGraph[[1]]$Nodes, Edges = tGraph[[1]]$Edges)
```

```{r, fig.height=7, fig.width=7}
plotData2D(Data = simple_tree, PrintGraph = tGraph[[1]] ,
           GroupsLab = rep(1, nrow(simple_tree)),
           Main = "A tree with 3 branches",
           Xlab = "Dimension 1", Ylab = "Dimension 2")
```


