---
title: "Projecting points on principal graphs"
author: "Luca Albergante (luca.albergante@gmail.com)"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

This document describes how to peoject points on a previously constructed principal graph.

## Building a tree

This example will show how to project points on a principal tree. However, the procedure is completely generic and can be used with any principal graph.

```{r}
library(rpgraph)
Data <- simple_tree
Results <- computeElasticPrincipalGraph(Data = Data, NumNodes = 30, Method = 'DefaultPrincipalTreeConfiguration')
```


By plotting the data in 2D we can see that different paths are present in the treee.

```{r, fig.height=7, fig.width=7}
plotData2D(Data = simple_tree, PrintGraph = Results[[1]] ,
           GroupsLab = rep(1, nrow(simple_circle)), Col = "black",
           Main = "Pincipal tree",
           Xlab = "Dimension 1", Ylab = "Dimension 2")
```

## Making a network

To get the different paths we can use graph analysis approach supported by the `igraph` library. Firts of all, we need to construct a network from the princial graph.

```{r}
library(igraph)
Net <- ConstructGraph(Results = Results[[1]], DirectionMat = NULL)
```

We can now obtain the end points by looking at nodes with degree 1 and find the shorthest path between them.

```{r}

EndPoints <- V(Net)[degree(Net) == 1]
AllPaths <- list()
for(StartIdx in 1:length(EndPoints)){
  Paths <- get.shortest.paths(graph = Net, from = EndPoints[StartIdx], to = EndPoints[-StartIdx])
  AllPaths <- append(AllPaths, Paths$vpath)
}

```

The paths found can be listed by looking at `AllPaths`.

```{r}
AllPaths
```


We can also look at their lengh distribution

```{r}
table(sapply(AllPaths, length))
```


```{r, fig.height=5, fig.width=5}
barplot(table(sapply(AllPaths, length)), main = "Paths between the endpoints", xlab = "length", ylab = "Frequency")
```

For this example, we will select one of the longest paths.

```{r}
SelPathIdx <- sample(which(sapply(AllPaths, length) == max(sapply(AllPaths, length))), 1)
SelPath <- AllPaths[[SelPathIdx]]
SelPath
```

## Projecting all the points

We can now project the points on the path

```{r}
TaxonList <- getTaxonMap(Results = Results[[1]], Data = Data)
Projections <- projectPoints(Results = Results[[1]], Data = Data, TaxonList = TaxonList, UseR = TRUE)
```

`Projections` will now contains the projections of the points on the graph in the `PointsOnEdgesCoords` field. We can use this information to confirm that the points are correctly projected on the tree.

```{r, fig.height=7, fig.width=7}
plot(Projections$PointsOnEdgesCoords[,1:2], xlab = "Dimension 1", ylab = "Dimension 2", main = "Point projections")
```


It is now possible to project the points on the selected path. Note that in the `igraph` representation of the principal graph the nodes have a "V_" prefix to their names that need to be removed when going back to the original point names which are defined by their names.

```{r}
NumericPath <- as.numeric(unlist(lapply(strsplit(SelPath$name, "V_"), "[[", 2)))
```

It is now possible to order the points on the selected path. Note that points associated with edges not present in the path will not considered.  

```{r}
PathProjection <- OrderOnPath(PrinGraph = Results[[1]], Path = NumericPath, PointProjections = Projections)
```

The ordered points can now be plotted and color-coded according to their position (i.e. `PathProjection$PositionOnPath`).

```{r, fig.height=7, fig.width=7}
plot(Projections$PointsOnEdgesCoords[order(PathProjection$PositionOnPath, na.last = NA),1:2], xlab = "Dimension 1", ylab = "Dimension 2", main = "Point projections", col=heat.colors(sum(!is.na(PathProjection$PositionOnPath))))
```
