---
title: "Using rpgraph to observe the construction of principal trees"
author: "Luca Albergante (luca.albergante@gmail.com)"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

This document describes how to produce a sequence of principal trees to that can be used to observe the dynamics of the algorithm and to produce animations.

## Building a principal tree

The `simple tree` dataset included in the package describes points placed on a three dimensional tree and we can use it to test the usage of principal trees. The `computeElasticPrincipalGraph` include an optional argument `NodeStep` that ca be used to obtain the intermediate steps. For example setting `NodeStep = 2` will save all the intermediate principal trees after 2 nodes have been added, i.e. will output tress with 2, 4, 6, 8, ... nodes.

```{r}
library(rpgraph)
Results <- computeElasticPrincipalGraph(Data = simple_tree, NumNodes = 20, Method = 'DefaultPrincipalTreeConfiguration', NodeStep = 2)
```

Each tree is saved as a different element of the `Results` list. `Results[[1]]` contains the tree with 2 nodes, `Results[[2]]` the tree with 4 nodes and so on. Therefore `Results` can be used to observe different snapshot of the construction process:

```{r, fig.height=10, fig.width=10}
par(mfcol = c(4,3))

for(i in 1:length(Results)){
  plotData2D(Data = simple_tree, PrintGraph = Results[[i]] ,
           GroupsLab = rep(1, nrow(simple_tree)),
           Main = paste("Pincipal tree", i),
           Xlab = "Dimension 1", Ylab = "Dimension 2")
}
```

or to produce a series of plot that can be used for an animation:

```{r, fig.height=7, fig.width=7}
for(i in 1:length(Results)){
  plotData2D(Data = simple_tree, PrintGraph = Results[[i]] ,
           GroupsLab = rep(1, nrow(simple_tree)),
           PlotProjections = "onEdges",
           Main = paste("Pincipal tree", i),
           Xlab = "Dimension 1", Ylab = "Dimension 2")
  Sys.sleep(1)
}
```





